Introduction to Structural Causal Models

### 2021-05-21 presentation_video.Rmd
graphical models

- Flexible simulations for higher order problems (intervention, counterfactual)
- Graphical visualization via directed acyclic graph

@@

- No need to specify exact parametric shape
- highlight colliders 
- estimation back door criterion 
-
- Nodes 
- Edges
- Parents/Ancestors/Descendents
- (Missing) Arrows

# Difference Path Diagramm and Pearls Causal Diagrams

@@

# References

@@

>>>>

- Flexible simulations for higher order problems (intervention, counterfactual)
- Graphical visualization via directed acyclic graph
- Example: SCM for fairness in dynamical system [@creager] : credit loan approval, time allocation, college admission

@@

- Nodes -> Variables (endogenous/exogenous)
- Edges -> relationship (equations)
- Parents/Ancestors/Descendents
- No need to specify exact parametric shape
- highlight colliders 
- estimation back door criterion 
- Test theoretical model structure via causal algorithms to detect structure in data (IC/PC Algo.)

@@

- Assosicational learning is easy to model because of lower information neccessary
- but not always appropiate in high stake settings 
- SCM as expressive simulator for causal modelling outperforms many other models
- computational advantage casting causal model as system of assignment equations 
- Enables modelling of higher order concepts like counterfactuals 
- Extensions through differential equations for concise modelling of time

\newpage 
# References

<div id="refs"></div>
\newpage

@@

\footnotesize

 |Method         | CBN |  SCM
|----------------|---------------------------|------------------------|
| Prediction     | $\boldsymbol{\cdot}$ Unstable                                                                                        | $\boldsymbol{\cdot}$ Stable                 
|                | $\boldsymbol{\cdot}$ Volatile to parameter changes                                                                   | $\boldsymbol{\cdot}$ More Natural Specification 
|                | $\boldsymbol{\cdot}$ Re-Estimate entire model                                                                        | $\boldsymbol{\cdot}$ Only estimate $\Delta$ CM
|                |                                                                                                                      | 
| Intervention   | $\boldsymbol{\cdot}$ Costly for Non-Markovian Models                                                                 | $\boldsymbol{\cdot}$ Pot. Cyclic Representation
|                | $\boldsymbol{\cdot}$ Unstable(Nature CP)                                                                             | $\boldsymbol{\cdot}$ Stable
|                | $\boldsymbol{\cdot}$ Only generic estimates($\Delta$ CP)                                                             | $\boldsymbol{\cdot}$ Context specific 
|                |                                                                                                                      | 
|Counterfactuals | $\boldsymbol{\cdot}$ **Impossible**                                                                                  | $\boldsymbol{\cdot}$ Possible
|                | $\boldsymbol{\cdot}$ no information on latent factors($\epsilon$)                                                    | $\boldsymbol{\cdot}$ Inclusion of latent factors
